\bf Landslide: \\ Systematic Dynamic Race Detection in Kernel-space

### 2012-05-18 design.tex
ryan's copyedit

Landslide's model for applying systematic exploration in kernel-space is comprised of four key points, which we overview here.

@@

	Landslide treats the sleep queue no differently from the runqueue when deciding which threads are runnable. The guest kernel will treat it differently, but since sleeping for a predetermined amount of time is never an appropriate way to solve race conditions, Landslide treats sleeping threads as notionally runnable. This fits directly into its model that runnable threads are ones that can be caused to run with a finite number of timer ticks in succession.
	\item {\bf No idling when progress can be made.} The kernel must not enter its idle loop (whether in an explicit idle thread or no) when the kernel is not truly idle. Landslide uses the idle loop to detect when a test begins/ends (test lifecycle tracking) and to detect when all threads in a test are wedged (bug detection).
\end{itemize}

@@

\begin{enumerate}
	\item {\bf Bootup.} Initial state. When the kernel is idling and the shell is waiting for keyboard input, advance state.
	\item {\bf Keyboard input.} Record the current thread count (the ``start population''). Generate keystrokes to cause the desired test case to run.
		When the current thread count is greater than the start population, advance state.
	\item {\bf Test running.} Begin constructing the decision tree here (since it is pointless to backtrack to before the test begins). When the current thread count equals the start population, advance state.
	\item {\bf Test cleanup.} The test processes have exited, but some housekeeping may yet remain. When the the kernel is idling and the shell is waiting for keyboard input, advance state.
	\item {\bf Test ended.} Invoke the explorer (Section~\ref{sec:components-explore}) to decide whom to rewind, and backtrack (to step 3).
\end{enumerate}

>>>>

Landslide's model for applying systematic exploration in kernel space is comprised of four key points, which we overview here.

@@

	Landslide treats the sleep queue no differently from the runqueue when deciding which threads are runnable. The guest kernel will treat it differently, but since sleeping for a predetermined amount of time is never an appropriate way to solve race conditions, Landslide treats sleeping threads as notionally runnable. This fits directly into its model that runnable threads are ones that can be caused to run with a finite number of timer ticks in succession.
	\item {\bf No idling when progress can be made.} The kernel must not enter its idle loop (whether in an explicit idle thread or not) when the kernel is not truly idle. Landslide uses the idle loop to detect when a test begins/ends (test lifecycle tracking) and to detect when all threads in a test are wedged (bug detection).
\end{itemize}

@@

\begin{enumerate}
	\item {\bf Bootup.} Initial state. When the kernel is idling and the shell is waiting for keyboard input, advance state.
	\item {\bf Keyboard input.} Record the current thread count (the ``start population''). Generate keystrokes to cause the desired test case to run.
		When the current thread count is greater than the start population, advance state.
	\item {\bf Test running.} Begin constructing the decision tree here (since it is pointless to backtrack to before the test begins). When the current thread count equals the start population, advance state.
	\item {\bf Test cleanup.} The test processes have exited, but some housekeeping may yet remain. When the the kernel is idling and the shell is waiting for keyboard input, advance state.
	\item {\bf Test ended.} Invoke the explorer (Section~\ref{sec:components-explore}) to decide where to rewind to, and backtrack (to step 3).
\end{enumerate}

### 2012-05-18 futurework.tex
ryan's copyedit

\begin{itemize}
	\item Landslide currently does not support ability to replay a particular choice sequence. If the user finds a bug, and wants to re-execute the interleaving that led to it, they have no choice but to re-explore the tree.
	\item Although Simics can pause execution and give the user a debug prompt, Landslide currently does not support being interrupted during exploration.
		This is because the current implementation uses a wrapper script around the Simics prompt which Landslide communicates with to perform backtracking.
		Though it would be difficult to redesign, it would be good to work around, to enable the user to interrupt, use the Simics debug prompt arbitrarily, and resume Landslide when ready.
	\item Landslide could print its decision traces in a better format. The current format is basically a ``raw data dump''. Raw text is probably the wrong interface; a graphical/tabular format with distinct columns for the execution of each different thread would be easier to understand.
	\item Landslide could attempt to detect when something has gone wrong that is not its fault (such as incorrect instrumentation, or the kernel violating some of its requirements), and print a useful error message rather than behaving mysteriously.
\end{itemize}

>>>>

\begin{itemize}
	\item {\bf Replay.} Landslide currently does not support ability to replay a particular choice sequence. If the user finds a bug, and wants to re-execute the interleaving that led to it, they have no choice but to re-explore the tree.
	\item {\bf Debug prompt support.} Although Simics can pause execution and give the user a debug prompt, Landslide currently does not support being interrupted during exploration.
		This is because the current implementation uses a wrapper script around the Simics prompt which Landslide communicates with to perform backtracking.
		Though it would be difficult to redesign, it would be good to work around, to enable the user to interrupt, use the Simics debug prompt arbitrarily, and resume Landslide when ready.
	\item {\bf Decision trace formatting.} Landslide could print its decision traces in a better format. The current format is basically a ``raw data dump''. Raw text is probably the wrong interface; a graphical/tabular format with distinct columns for the execution of each different thread would be easier to understand.
	\item {\bf Feedback for incorrect instrumentation.} Landslide could attempt to detect when something has gone wrong that is not its fault (such as incorrect instrumentation, or the kernel violating some of its requirements), and print a useful error message rather than behaving mysteriously.
\end{itemize}

### 2012-05-18 related.tex
ryan's copyedit

DataCollider~\cite{datacollider} is a tool for detecting data races, a special kind of concurrency bug which involve concurrent unprotected accesses to memory that can interleave unsafely on a fine-grained level.
DataCollider uses memory access sampling, and remains largely agnostic of synchronisation operations, to find data races in the Windows 7 kernel.
In Section~\ref{sec:future-new} we discuss how systematic execution might be enhanced by integrating data race detection techniques.

Carburizer~\cite{carburizer} is an effort to make imperfect kernel-level drivers work more reliably in the presence of faulty peripheral devices. They use static analysis to identify variables ``tainted'' by device input and unsafe uses thereof. Carburizer is able to transform such badly-written code to be more robust.
Kernel device drivers are a ripe breeding ground for bugs, because they are frequently written by less careful developers, and the sheer volume of code makes it difficult to rigorously review.
Due to the interrupt-driven nature of many drivers, they are also prone to concurrency bugs, which Carburizer does not address.

>>>>

DataCollider~\cite{datacollider} is a tool for detecting data races, a special kind of concurrency bug which involves concurrent unprotected accesses to memory that can interleave unsafely on a fine-grained level.
DataCollider uses memory access sampling, and remains largely agnostic of synchronisation operations, to find data races in the Windows 7 kernel.
In Section~\ref{sec:future-new} we discuss how systematic execution might be enhanced by integrating data race detection techniques.

Carburizer~\cite{carburizer} is an effort to make imperfect kernel-level drivers work more reliably in the presence of faulty peripheral devices. They use static analysis to identify variables ``tainted'' by device input and unsafe uses thereof. Carburizer is able to transform such badly-written code to be more robust.
Kernel device drivers are a ripe breeding ground for bugs, because they are frequently written by less careful developers, and the sheer volume of code makes them difficult to rigorously review.
Due to the interrupt-driven nature of many drivers, they are also prone to concurrency bugs, which Carburizer does not address.

### 2012-05-18 terminology.tex
ryan's copyedit

		Figure~\ref{fig:threadfork} depicts a decision tree for a simple race with three possible interleavings.
	\item {\bf Branch}:
		One particular execution of the test case; a set of decision points with exactly one transition between each of them, characterising a single interleaving of threads.
	\item {\bf Interleaving}:
		A less precise / more abstract term for a branch, or for a subset of transitions which make up a branch.
	\item {\bf Bug}: An execution state which Landslide identifies as incorrect, which is indicative of illegal behaviour. Each branch of the tree may or may not have a bug in it. See Section~\ref{sec:techniques-bugs}.
	\item {\bf Backtracking}:
		Performed at the end of each branch of the tree, when the test case has finished running. Landslide identifies a decision point from the current branch at which the \revision{next interleaving to explore diverges from the current one}, reverts the state of the guest kernel to what it was at that decision point, and causes a different thread to run, hence exploring a different branch/interleaving.
	\item {\bf Decision trace}:
		A list of information about all decision points in a given branch, printed when a bug is found to help the user analyse its cause. Contains, for each decision point, the TID of the thread that was running previously, the TID that was newly chosen, the current instruction pointer, and the stack trace of the previous thread at the point it was switched away from.\footnote{Though Landslide does not provide it, a decision trace could also contain for each transition a list of shared memory accesses that conflicted with other transitions, and/or a listing of happens-before relationships between transitions.}
\end{enumerate}

>>>>

		Figure~\ref{fig:threadfork} depicts a decision tree for a simple race with three possible interleavings.
	\item {\bf Branch}:
		One particular execution of the test case; a set of decision points with exactly one transition between each of them, characterising a single interleaving of threads.
	\item {\bf Interleaving}:
		A less precise / more abstract term for a branch, or for a subset of transitions which make up a branch.
	\item {\bf Bug}: An execution state which Landslide identifies as incorrect, which is indicative of illegal behaviour. Each branch of the tree may or may not have a bug in it. See Section~\ref{sec:techniques-bugs}.
	\item {\bf Backtracking}:
		Performed at the end of each branch of the tree, when the test case has finished running. Landslide identifies a decision point from the current branch at which the \revision{next interleaving to explore diverges from the current one}, reverts the state of the guest kernel to what it was at that decision point, and causes a different thread to run, hence exploring a different branch/interleaving.
	\item {\bf Decision trace}:
		A list of information about all decision points in a given branch, printed when a bug is found to help the user analyse its cause. The trace contains, for each decision point, the TID of the thread that was running previously, the TID that was newly chosen, the current instruction pointer, and the stack trace of the previous thread at the point it was switched away from.\footnote{Though Landslide does not provide it, a decision trace could also contain for each transition a list of shared memory accesses that conflicted with other transitions, and/or a listing of happens-before relationships between transitions.}
\end{enumerate}

### 2012-05-18 thesis.tex
ryan's copyedit

In kernel space, race condition debugging becomes even more difficult. Many aspects of the concurrency implementation itself are part of the system being tested. While user-space testing frameworks rely on the underlying kernel to provide common concurrency abstractions which the framework can use to drive the test, systematic exploration in the kernel itself necessarily exists underneath the level of such abstractions.
Furthermore, a tool that supports testing multiple kernels must support many different possible scheduler designs.

>>>>

In kernel space, race condition debugging becomes even more difficult. Many aspects of the concurrency implementation itself are part of the system being tested. While user-space testing frameworks rely on the underlying kernel to provide common concurrency abstractions which the framework can use to drive the test, systematic exploration in the kernel itself necessarily exists underneath such abstractions.
Furthermore, a tool that supports testing multiple kernels must support many different possible scheduler designs.
